\label{sec:preface}

\begin{comment}
Foreword / Preface / Acknowledgements
This is the place for "informal chat" like how the master thesis task has progressed and acknowledgment to those who one feels for thanking. One can also include what kind of software that has been used in the master thesis work. Some choose to have the foreword before the Table of Contents (TOC).
\end{comment}
% Takk til veiledere
I want to thank my supervisors, Andrea Storås, Michael Riegler, and Kyrre Glette, for all your help, motivation, and interesting discussions.
I am thankful that it was you guys that followed me on this journey – without your encouraging presence, this project would not be the same.

% Takk til folk og ML-nodene 
All the experiments conducted during this work have been run on the machine learning nodes at the AI HUB hosted at the University of Oslo. These are now becoming a part of the Norwegian AI Cloud (NAIC). 
Therefore, I would like to thank these nodes for their service, and may you have a prosperous life serving new and interesting assignments.

% Takk til AI-generert forsidebilde
The header image of the cute robot on the front page is made with the help of the generative text-to-image model Stable Diffusion. The image is meant to illustrate a large language model, reading lots of books to gain knowledge. It is wonderful to live in a time where a stressed student can make a beautiful image without having any artistic qualities.


% Hva jeg har lært
This thesis has introduced me to the depths of a very fascinating field of research.
The field of AI and XAI is developing at an extraordinary speed, which is both exciting and important to be aware of, and I am curious to follow the coming advancements. 
Open-sourced AI methods have been essential during the development of this work.

Astonishing advancements are now made in both AI and XAI, and their discoveries help contribute to further democratizing these helpful tools. It is beneficial for interested individuals like me to have models and documentation openly available, and it contributes to the transparency of the extreme advancement.

% Takk til venner og familie og Vics
I want to give a special thanks to all the lovely people at Blindern Studenterhjem. You have given me the best student experience that I could ever dream of. 
Last but certainly not least, I would like to give a huge printed hug to my family and Victoria for motivating me and supporting me throughout this entire voyage.