\label{sec:2_background_all_content}

This chapter provides an overview of the background knowledge and related work needed for understanding the research performed in this thesis. The chapter begins by explaining the motivation for this research and the importance of \gls{xai}. 
The chapter also presents existing technologies relevant to the research, such as \gls{ai}, machine learning approaches, and model explanation techniques. It discusses the current state-of-the-art in the \gls{xai} field, including recent advancements and research trends. Additionally, the chapter summarizes the most relevant evaluation metrics, which are essential for assessing the effectiveness of \gls{ai} and \gls{xai} techniques.

Finally, this chapter provides an overview of the related work done in the \gls{xai} field, focusing on the most relevant studies and research findings. It examines the approaches and techniques used in these studies and discusses their strengths and limitations. 
The chapter then discusses the key concepts and principles of \gls{xai}, including the various types of interpretability and the methods used to achieve it.

In essence, this chapter lays the foundation for the subsequent chapters of this thesis, providing the necessary background and context for understanding the research presented.

\input{section/2_background/2_background_problem_and_application}
\input{section/2_background/2_background_theory.tex}
\input{section/2_background/2_related_work}
\input{section/2_background/2_summary}